Dynamic Species Interactions in the San Nicolas Island Kelp Forest

### 2018-07-18 EDM_SNI_analysis.Rmd
fixed physical normalization, added density plots, new network graph

```{r setup, include=FALSE,message=F}
# Load required packages
require(tidyverse)
require(rEDM)
require(quantreg)
require(knitr)
require(kableExtra)
require(extrafont)

@@

```{r ccm example}
tempE <- westend.bestE['purp']
temp <- ccm(westend,lib=as.matrix(westend.segs),pred=as.matrix(westend.segs),E=tempE,lib_column= 'mac',target_column = 'purp',lib_sizes = c(10,25,50,75,100,125,150,200,300,400,500),num_samples=100,replace=T,silent=T,RNGseed = 41389)

mac_xmap_purp <- temp %>%
  group_by(lib_size)%>%
  summarise(rhomean=mean(rho,na.rm=T),upper=rhomean+sd(rho,na.rm=T),lower=rhomean-sd(rho,na.rm=T))%>%
  ungroup()%>%
  ggplot(aes(lib_size,rhomean))+
  geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
  geom_line(color="darkorchid3")+
  labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title="Macrocystis xmap Purple Urchin")

mac_xmap_purp
# ggsave("plots/ccm_example.png",plot=mac_xmap_purp,width=6,height=4)
```

@@

      # second p-value (rho at lib-size 500 greater than rho at lib-size 10? By more than 0.1?)
      p2 <- t.test(temp$rho[temp$lib_size==10],temp$rho[temp$lib_size==500])$p.value
      if((rhomeans$rho[rhomeans$lib_size==500]-rhomeans$rho[rhomeans$lib_size==10])<0.1) p2 <- 0
      westend.p2.mat[i,j] <- as.numeric(p2)

@@

```{r igraph, echo=F, message=F,warning=F,fig.height=6}
require(igraph) # R Package for network graphing and analysis

# Add NAs for connections in to physical variables (since we want them just to be exogenous)
connect.mat <- t(westend.ptot.mat) %>% 
  as.data.frame() %>%
  mutate(mei=rep(NA,nrow(.)),pdo=rep(NA,nrow(.)),npgo=rep(NA,nrow(.)),waves=rep(NA,nrow(.)),sst=rep(NA,nrow(.)))

g <- graph.adjacency(as.matrix(connect.mat))
par(mfrow=c(1,1))

# Variable names

names <- as.character(sapply(V(g)$name, function(x) namekey$plotting[match(x,namekey$short)]))
names [7:11] <- c("MEI","PDO","NPGO","Max Hs","SST")

# Vertex labels
V(g)$name <-names
vert_in <- sapply(degree(g,mode="in"),function(x) paste0("(",x,")"))
vert_out_in <- paste0(degree(g,mode="out"),vert_in)
V(g)$label <- paste(sep="\n",V(g)$name, vert_out_in)
V(g)$label[7:11] <- paste(V(g)$name[7:11])

#vertex coordinates
crds <- cbind(c(116.0000, 318.0000, 177.0000, 581.0000, 346.0000, 385.0000, 331.1627, 151.8379, 573.0000, 539.0000,  45.0000),c(0.0000, 186.0000, 362.0000, 172.0000,  40.0000, 374.0000, 537.0000, 504.2360, 35.0000, 406.0000, 288.4111))

# vector of node colors for plotting
vert.cols <- c(rep("tan3",3),rep("orchid3",2),"tan3",rep("gray70",5))
plot(g,asp=0.8, vertex.label.color="black", vertex.label.cex=0.8, edge.arrow.size=0.6,vertex.label.font=2, vertex.shape="circle", edge.curved=F,margin=0, vertex.size=30,rescale=T,edge.lty=1,edge.color="gray40",layout=crds,vertex.color=vert.cols,vertex.frame.color=NA,main="")
```

@@

  # box plot of interactions
  sppname <- namekey$long[match(species,namekey$short)]
  int.box <- ggplot(coeffs.long,aes(x=plotting,y=value))+
    geom_boxplot() +
    geom_hline(yintercept=0,linetype=2,col="gray40")+
    xlab("Interacting Variable") +
    ylab("Coefficient") +
    ggtitle(paste("Interactive Effects on Dynamics of ",sppname))

@@

all.coeffs.with.phys<- westend %>%
  select(one_of(c(phys.vars,"period"))) %>%
  distinct() %>%
  right_join(spp.only.mod.coeffs,by="period")%>%
  gather("external_forcing","ext_value",one_of(phys.vars))

@@

```{r individual_effects,fig.width=7}
phys_effects_on_interactions <- purrr::map(as.list(study_spp),function(spp) {
  quants <- c(0.05,0.5,0.95)
  plot_title <- namekey$long[match(spp,namekey$short)]
  all.coeffs.with.phys %>%
  filter(!is.na(type),mod==spp)%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
    geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
    geom_point(col='gray60',alpha=0.6)+
    geom_hline(yintercept = 0,linetype=2)+
    geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
    scale_color_manual(values=c('mac'="navyblue",'ymac'="gray30",'pter'="darkgreen",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),name="effect")+
    labs(title=paste("Physical forcing effect on interactions with",plot_title),x="Normalized Index Value",y="Interaction Strength")+
    facet_grid(type~external_forcing,scales = "free_y")+
    theme(strip.text=element_text(size=8),
          legend.text = element_text(size=12),
          legend.position="right",
          legend.margin=margin(t = -0.5, unit='line'),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line")
        )
})
ggsave("plots/phys_effects_red.png",plot=phys_effects_on_interactions[[1]])
ggsave("plots/phys_effects_purp.png",plot=phys_effects_on_interactions[[2]])
ggsave("plots/phys_effects_lam.png",plot=phys_effects_on_interactions[[3]])
ggsave("plots/phys_effects_pter.png",plot=phys_effects_on_interactions[[4]])
ggsave("plots/phys_effects_mac.png",plot=phys_effects_on_interactions[[5]])
ggsave("plots/phys_effects_ymac.png",plot=phys_effects_on_interactions[[6]])

@@

```{r some single plots}
# macrocystis, pterygophora, and temperature
quants<-c(0.05,0.5,0.95)
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('ymac','pter'),effect %in% c('mac','pter'),external_forcing=="sst")%>%

  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
    geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
    geom_point(col='black',alpha=0.6)+
    geom_hline(yintercept = 0,linetype=2,col="black")+
    geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
    scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
    labs(x="Normalized SST",y="Interaction Strength")+
    facet_grid(~mod,scales = "free_y")+
    theme(plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and NPGO
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
  geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+ 
  geom_point(col='black',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="black")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  theme(plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, herbivory, and NPGO
all.coeffs.with.phys %>%
  filter(type=="herbivory",mod%in% c('mac','pter'),effect %in% c('red','purp'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
    geom_point(col='black',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="black")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
    theme(plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and waves
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="waves")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
  geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
  geom_point(col='black',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="black")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized SWH",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

@@

### Effect of disturbance on *Macrocystis*
```{r disturbance and mac,fig.height=7,fig.width=5}
quants <- c(0.05,0.5,0.95)
waves.mac.pter <- westend %>%
  select(one_of(c("pter","lam","mac","period"))) %>%
  distinct() %>%
  right_join(all.mod.coeffs.long,by="period")%>%
  gather("external_spp","ext_value",one_of(c('pter','lam','mac')))%>%
  filter(effect%in% c('waves','pter','lam'),mod=='mac')%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Density",y="Effect of SWH on Macrocystis")+
  facet_grid(external_spp~effect,scales="free")
waves.mac.pter
```

### Effect of Macrocystis on Laminaria/Pterygophora interaction
```{r disturbance and mac,fig.height=7,fig.width=5}
quants <- c(0.05,0.5,0.95)
mac.lam.pter <- westend %>%
  select(one_of(c("mac","period"))) %>%
  distinct() %>%
  right_join(all.mod.coeffs.long,by="period")%>%
  gather("external_spp","ext_value",mac)%>%
  filter(effect=='pter',mod=='lam')%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Density",y="Effect of Pterygophora on Laminaria")
mac.lam.pter
```
### Effect of urchin density on Macrocystis/Pterygophora interaction

```{r urchin dens mac pter}
quants <- c(0.05,0.5,0.95)
urchin.mac.pter <- westend %>%
  select(one_of(c("red","period"))) %>%
  distinct() %>%
  mutate(time=row_number())%>%
  right_join(all.mod.coeffs.long,by=c("period",'time'))%>%
  gather("external_spp","ext_value",red)%>%
  filter(effect %in% c('pter','mac'),mod %in% c('pter','mac'),!is.na(ext_value))%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Red Urchin Density",y="Interaction Strength")+
  facet_wrap(~mod)
urchin.mac.pter
```

### Do conditions affect average interaction strength?

```{r conditions and interaction strength}
# disturbance. top quartile vs. bottom quartile
all.coeffs.with.phys %>%
  mutate(av_value=abs(value))%>%
  filter(external_forcing=='waves', type=='algal competition', ext_value>=quantile(westend$pdo,0.75,na.rm=T)|ext_value<=quantile(westend$pdo,0.25,na.rm=T))%>%
  mutate(stress_level=ifelse(ext_value>=quantile(westend$pdo,0.75,na.rm=T),'high','low'))%>%
  ggplot(aes(log(value),..density..,fill=stress_level))+
  geom_density(alpha=0.6)+
  geom_vline(xintercept = 0,linetype=2,color='black')+
  labs(x='Interaction Strength',y='Probability Density')

@@

```{r ints block}
# create time series block of all interactions (later we could come back to this to deal with error)
interactions.block <- spp.only.mod.coeffs %>%
  filter(type != "intraspecies", !is.na(type))%>%
  select(time,period,mod,effect,value)%>%
  unite("effect_mod",effect,mod)%>%
  spread(effect_mod,value) %>%
  complete(time=full_seq(x=c(1,630),1))%>%
  # join physical variables
  left_join(select(westend,period,one_of(phys.vars)),by="period")%>%
  distinct()
```

```{r simplex ints}
# Run simplex projection for each species interaction
interaction_names <- colnames(interactions.block)[3:29]
interactions.simp.list <- list()

@@

# Save best embedding dimensions
ints.bestE <- purrr::map_chr(interactions.simp.list,function(x) {
  temp <- x %>% filter(!is.na(rho))
  temp$E[temp$rho==max(temp$rho)]
})
```

@@

## Run CCM for each combination of variables
for(i in 1:n_row) {
  for(j in 1:n_col) {
      # remember, we use the best embedding dimension for the target variable (the variable we're cross-mapping to, i.e. the putative forcing variable, in this case the physical variables)
      tempE= westend.bestE[col_names[j]]
      temp <- ccm(interactions.block,lib=as.matrix(westend.segs),pred=as.matrix(westend.segs),E=tempE,lib_column= 2+i,target_column = 27+j,
                  lib_sizes = c(10,500),num_samples=100,replace=T,silent=T,RNGseed = 41389)
      # mean rho at library size 500
      rhomeans <- temp %>% ccm_means()
      interactions.xmap_mat[i,j] <- rhomeans$rho[rhomeans$lib_size==500]

@@

## plot
interactions.xmap_all_plot <- ggplot(interactions.xmap_rast,aes(x=force.name,y=predictee,fill=rho)) +
  geom_raster() +
  scale_fill_gradient(low = "#9AFF9A", high = "#548B54", space = "Lab", na.value = "grey50", 
      guide = "colourbar",limits=c(0,0.8),breaks=c(0.2,0.4,0.6,0.8),name=expression(paste(rho, "(skill)"))) +
  ggtitle("Interaction Convergent Cross Mapping") +
  xlab("Predictor (Forcing Variable)") +
  ylab("Predicted Variable") +
  theme(text=element_text(color="black"),
        plot.background = element_rect(fill="white"),
        axis.text=element_text(color="black"),
    axis.text.x=element_text(angle = 90, hjust = 1,vjust=0.1),
    panel.border = element_blank())

@@

>>>>

```{r setup, include=FALSE,message=F}
# Load required packages
require(tidyverse)
require(rEDM)
require(quantreg)
require(knitr)
require(kableExtra)
require(extrafont)
require(GGally)
require(sna)
require(network)

@@

```{r ccm example}
tempE <- westend.bestE['red']
temp <- ccm(westend,lib=as.matrix(westend.segs),pred=as.matrix(westend.segs),E=tempE,lib_column= 'mac',target_column = 'red',lib_sizes = c(10,25,50,75,100,125,150,200,300,400,500),num_samples=100,replace=T,silent=T,RNGseed = 41389)

mac_xmap_red <- temp %>%
  group_by(lib_size)%>%
  summarise(rhomean=mean(rho,na.rm=T),upper=quantile(rho, 0.975),lower=quantile(rho, 0.025))%>%
  ungroup()%>%
  ggplot(aes(lib_size,rhomean))+
  geom_ribbon(aes(ymin=lower,ymax=upper),alpha=0.3,fill="red")+
  geom_line(color="darkorchid3")+
  labs(x="Library Size",y=expression(paste(rho, " (predictive skill)")),title="Macrocystis xmap Red Urchin")

mac_xmap_red
# ggsave("plots/ccm_example.png",plot=mac_xmap_purp,width=6,height=4)
```

@@

      # second p-value (rho at lib-size 500 greater than rho at lib-size 10? By more than 0.1?)
      p2 <- t.test(temp$rho[temp$lib_size==10],temp$rho[temp$lib_size==500])$p.value
      if((rhomeans$rho[rhomeans$lib_size==500]-rhomeans$rho[rhomeans$lib_size==10])<0.1) p2 <- 1
      westend.p2.mat[i,j] <- as.numeric(p2)

@@

```{r network,fig.height=6,fig.width=6}
connect.mat <- t(westend.ptot.mat) %>% 
  as.data.frame() %>%
  mutate(mei=rep(0,nrow(.)),pdo=rep(0,nrow(.)),npgo=rep(0,nrow(.)),waves=rep(0,nrow(.)),sst=rep(0,nrow(.)))%>%
  as.matrix()
names <- map_chr(colnames(connect.mat), function(x) as.character(namekey$plotting[match(x,namekey$short)]))
names [7:11] <- c("MEI","PDO","NPGO","Max Hs","SST")
g <- network(connect.mat,vertex.attrnames = names,loops=F)
g %v% "type" <- c(rep("darkgreen",3),rep("darkorchid1",2),"darkgreen",rep("dodgerblue",5))
g %v% "labels" <- names
int_net<-ggnet2(g,arrow.size=8,arrow.gap=0.05,color = "type")+
  geom_point(aes(color = color), size = 12, alpha = 0.5) +
  geom_point(aes(color = color), size = 9)+
  geom_text(aes(label=label),family="Rockwell",size=4)+
  theme(panel.border = element_rect(colour="white",fill=NA))

ggsave("plots/int_net.png",plot=int_net)
int_net
```

@@

  # box plot of interactions
  sppname <- species
  int.box <- coeffs.long %>% filter(effect!='const')%>%
    ggplot(aes(x=plotting,y=value))+
    geom_boxplot() +
    geom_hline(yintercept=0,linetype=2,col="gray40")+
    xlab("Interacting Variable") +
    ylab("Coefficient") +
    ggtitle(paste("Interactive Effects on Dynamics of ",sppname))

@@

spp.only.coeffs.with.phys<- westend %>%
  select(one_of(c(phys.vars,"period"))) %>%
  distinct() %>%
  right_join(filter(spp.only.mod.coeffs,!is.na(type)),by="period")%>%
  gather("external_forcing","ext_value",one_of(phys.vars))

@@

```{r individual_effects,fig.width=7}
phys_effects_on_interactions <- purrr::map(as.list(study_spp),function(spp) {
  quants <- c(0.05,0.5,0.95)
  plot_title <- namekey$long[match(spp,namekey$short)]
  spp.only.coeffs.with.phys %>%
  filter(!is.na(type),mod==spp)%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
    # geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
    geom_point(col='gray60',alpha=0.6)+
    geom_hline(yintercept = 0,linetype=2)+
    geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
    scale_color_manual(values=c('mac'="navyblue",'ymac'="gray30",'pter'="darkgreen",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),name="effect")+
    labs(title=paste("Physical forcing effect on interactions with",plot_title),x="Normalized Index Value",y="Interaction Strength")+
    facet_grid(type~external_forcing,scales = "free_y")+
    theme(strip.text=element_text(size=8),
          legend.text = element_text(size=12),
          legend.position="right",
          legend.margin=margin(t = -0.5, unit='line'),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line")
        )
})
ggsave("plots/phys_effects_red.png",plot=phys_effects_on_interactions[[1]])
ggsave("plots/phys_effects_purp.png",plot=phys_effects_on_interactions[[2]])
ggsave("plots/phys_effects_lam.png",plot=phys_effects_on_interactions[[3]])
ggsave("plots/phys_effects_pter.png",plot=phys_effects_on_interactions[[4]])
ggsave("plots/phys_effects_mac.png",plot=phys_effects_on_interactions[[5]])
ggsave("plots/phys_effects_ymac.png",plot=phys_effects_on_interactions[[6]])

@@

@@

### Example: Effect of Macrocystis on Pterygophora

```{r mac pter}
mac_effect_on_pter <-spp.only.coeffs.with.phys %>%
  ungroup()%>%
  filter(mod=="pter",effect=="mac")%>%
  mutate(phys_high_low=case_when(
    ext_value > 1 ~ "high",
    ext_value < -1 ~ "low")) %>%
  filter(!is.na(phys_high_low))%>%
  left_join(select(westend,period,one_of(phys.vars)),by="period")%>%
  group_by(type,external_forcing,phys_high_low)%>%
  mutate(mean_int=mean(value,na.rm=T))%>%
  distinct()%>%
  ungroup()

mac_effect_on_pter %>%
ggplot(aes(value,..density..,fill=phys_high_low))+
  geom_density(col=NA,alpha=0.6)+
  geom_vline(xintercept=0,linetype=2)+
  geom_vline(aes(xintercept=mean_int,color=phys_high_low))+
  scale_fill_manual(values=c("red","blue"),name="Physical\nForcing Value")+
  scale_color_manual(values=c("red","blue"))+
  guides(color="none")+
  labs(x="Effect of Macrocystis on Pterygophora",y="Probability Density")+
  facet_grid(rows=vars(external_forcing))

### Distribution of interaction strengths by type

```{r interactions density plot by type,fig.height=6}
# Density plot
interaction_distribution_by_type <- spp.only.mod.coeffs %>%
  filter(!is.na(type))%>%
  group_by(type) %>%
  mutate(mean_int=mean(value,na.rm=T))%>%
  ungroup()%>%
  ggplot(aes(value,..density..))+
  geom_density(aes(fill=type),alpha=0.8,col=NA)+
  geom_vline(xintercept=0,linetype=2)+
  geom_vline(aes(xintercept=mean_int))+
  scale_fill_manual(values=c("darkorange1","orangered2","darkorchid1","palegreen3","skyblue1"),name="")+
   guides(fill=F)+
  xlim(-2,2)+
  labs(x="Interaction strength",y="Kernel density")+
  geom_text(aes(x=-2,y=0.5,label=type),hjust="inward",family="Rockwell",fontface="italic",check_overlap = TRUE,size=4,show.legend = FALSE)+
  facet_grid(rows=vars(type))+
  theme(strip.text = element_blank())

interaction_distribution_by_type
ggsave("plots/int_dist_by_type_density.png",plot=interaction_distribution_by_type,h=6)
```

```{r ints dist by type and phys}
int_dist_by_type_phys <- spp.only.coeffs.with.phys %>%
  ungroup()%>%
  mutate(phys_high_low=case_when(
    ext_value > 1 ~ "high",
    ext_value < -1 ~ "low")) %>%
  left_join(select(westend,period,one_of(phys.vars)),by="period")%>%
  group_by(type,external_forcing,phys_high_low)%>%
  mutate(mean_int=mean(value,na.rm=T))%>%
  distinct()%>%
  ungroup()

int_dist_by_type_phys_plots <- purrr::map(phys.vars,function(x) {
  int_dist_by_type_phys %>%
    filter(external_forcing==x,!is.na(phys_high_low))%>%
    ggplot(aes(value,..density..,fill=phys_high_low,col=phys_high_low))+
    geom_density(alpha=0.6,col=NA)+
    geom_vline(xintercept=0,linetype=2)+
    geom_vline(aes(xintercept=mean_int,col=phys_high_low))+
    xlim(-2,2)+
    scale_fill_manual(values=c("red","blue"),name=paste0(x,"\nLevel"),labels=c("high (> 1)","low (< 1)"))+
    scale_color_manual(values=c('red','blue'))+
    guides(color="none")+
    geom_text(aes(x=-2,y=0.5,label=type),col="black",hjust="inward",family="Rockwell",fontface="italic",check_overlap = TRUE,size=4,show.legend = FALSE)+
    labs(x="Interaction strength",y="Kernel density")+
    facet_grid(rows=vars(type))+
    theme(strip.text=element_blank())
  })
names(int_dist_by_type_phys_plots) <- phys.vars
int_dist_by_type_phys_plots$mei
int_dist_by_type_phys_plots$sst
int_dist_by_type_phys_plots$pdo
int_dist_by_type_phys_plots$waves
int_dist_by_type_phys_plots$npgo

@@

```{r ints block}
# create time series block of all interactions (later we could come back to this to deal with error)
interactions.block <- spp.only.mod.coeffs %>%
  filter(type != "intraspecies", !is.na(type))%>%
  select(time,period,mod,effect,value)%>%
  unite("effect_mod",effect,mod)%>%
  spread(effect_mod,value) %>%
  complete(time=full_seq(x=c(1,630),1))%>%
  # join physical variables
  left_join(select(westend,period,one_of(phys.vars)),by="period")%>%
  distinct()%>%
  ungroup()%>%
  # normalize
  mutate_at(vars(lam_mac:ymac_red),function(dens) {(dens-mean(dens,na.rm=T))/sd(dens,na.rm=T)})
```

```{r ints time series plot}
spp.only.mod.coeffs %>%
  filter(time %in% 1:63,type=="algal competition")%>%
  unite("interaction",effect,mod,sep="_")%>%
  ggplot(aes(time,value,color=interaction))+
  geom_line()+
  geom_point()+
  labs(x="period",y="interaction strength")
```

```{r simplex ints}
# Run simplex projection for each species interaction
interaction_names <- colnames(interactions.block)[3:22]
interactions.simp.list <- list()

@@

# Save best embedding dimensions
ints.bestE <- purrr::map_chr(interactions.simp.list,function(x) {
  temp <- x %>% filter(!is.na(rho))
  temp$E[temp$rho==max(temp$rho)]
})
purrr::map_chr(interactions.simp.list,function(x) {
  max(x$rho,na.rm=T)
})
```

@@

## Run CCM for each combination of variables
for(i in 1:n_row) {
  for(j in 1:n_col) {
      # remember, we use the best embedding dimension for the target variable (the variable we're cross-mapping to, i.e. the putative forcing variable, in this case the physical variables)
      tempE= westend.bestE[col_names[j]]
      temp <- ccm(interactions.block,lib=as.matrix(westend.segs),pred=as.matrix(westend.segs),E=tempE,lib_column= 2+i,target_column = 2+length(interaction_names)+j,
                  lib_sizes = c(10,500),num_samples=100,replace=T,silent=T,RNGseed = 41389)
      # mean rho at library size 500
      rhomeans <- temp %>% ccm_means()
      interactions.xmap_mat[i,j] <- rhomeans$rho[rhomeans$lib_size==500]

@@

## plot
interactions.xmap_all_plot <- ggplot(interactions.xmap_rast,aes(x=force.name,y=predictee,fill=rho)) +
  geom_raster() +
  scale_fill_gradient(low = "#9AFF9A", high = "#548B54", space = "Lab", na.value = "grey50", 
      guide = "colourbar",limits=c(0,0.75),breaks=c(0.25,0.5,0.75),name=expression(paste(rho, "(skill)"))) +
  ggtitle("Interaction Convergent Cross Mapping") +
  xlab("Predictor (Forcing Variable)") +
  ylab("Predicted Variable") +
  theme(text=element_text(color="black"),
        plot.background = element_rect(fill="white"),
        axis.text=element_text(color="black"),
    axis.text.x=element_text(angle = 90, hjust = 1,vjust=0.1),
    panel.border = element_blank())

@@

### Multivariate S-maps with Interactions

```{r multivariate smaps}
ints.causal.vars <- purrr::map(interaction_names, function(x) {
  interactions.xmap_rast %>%
    filter(predictee==x,!is.na(rho))->out
  out <- out$predictor
  out
})

names(ints.causal.vars)=interaction_names

# apply s-maps
ints.multi.mods <- purrr::map(interaction_names,function(x) {
  if(length(ints.causal.vars[[x]])==0) return(NULL)
  else(smap_multi(interactions.block,westend.segs,x,ints.causal.vars[[x]]))
})
names(ints.multi.mods) <- interaction_names

purrr::map_dbl(ints.multi.mods, function(x) {
x$model$rho
})
